Clustering of Last.fm Music

### 2015-05-07 writeup/MSDgroupwriteup.tex
Table
Update MSDgroupwriteup.tex

\begin{abstract}
This is the abstract. 
\end{abstract}

@@

\section{Introduction}
In the online world, recommendation systems are used in shopping, entertainment and even dating. 
Because of their popularity and usefulness, there are numerous incentives to implement a good recommendation system. 
In the music industry, companies use recommendation systems to provide a better music streaming radio service to their listeners. 
With the music industry shifting away from a more traditional distribution model of physical CDs or records, 
having an efficient and well performing recommendation system is very important.

In building a recommendation system, there are many problems to tackle. This is why some methods take more information about the user and the song into account.
There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.

For our project, we chose to explore various song recommendation systems algorithms such as matrix factorization, non-negative matrix factorization as well as user based and item based similarity measures.  
% memory based collaborative filtering

@@

\subsection{User Based and Item Based Collaborative Filtering}
%\subsubsection{Intro to Memory-Based Collaborative Filtering}

A simpler set of methods for determining which songs to recommend to individual users is by using user- or item/song- based similarity (both also considered collaborative filtering). These only consider whether or not a user listened to a song and does not take into account the play count. Similarity scores between users (i.e. user-based) and songs (i.e. item-based) were calculated using cosine similarity. Intuitively, they work by recommending users songs that other similar users like (user-based) or by recommending songs which are similar to the songs the user has already listened to (item-based).

\subsubsection{User-Based Collaborative Filtering}

For user-based recommendation, we first calculated the similarity score between every pair of users, $u$ and $v$, using the equation below:

@@

\subsubsection{Item-Based Collaborative Filtering}

For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, $i$ and $j$, using the equation below and saved the most similar songs to each song in the process:

@@

\section{Conclusion}

>>>>

\begin{abstract}
We attempt to build a recommendation system for a subset of the Million Song Dataset. We explore various algorithms including matrix factorization, user-based collaborative filtering, and item-based collaborative filtering. In each case we recommend the top 500 songs that our algorithm returns to each user and compare against the testing set of what the user actually listened to. Matrix factorization gives the worst results, only slightly above the popularity baseline, while artist-based popularity and user-based and item-based collaborative filtering methods yield better results.
\end{abstract}

@@

\section{Introduction}
In the online world, recommendation systems are becoming increasingly important in various industries including in retail, entertainment, and even dating. 
Because of their popularity and usefulness, there are numerous incentives to implement a good recommendation system. 
In the music industry, companies use recommendation systems to provide a better service to their listeners. 
With the music industry shifting away from a more traditional distribution model of physical CDs or records, 
having an efficient and well performing recommendation system is very important.

In building a recommendation system, we had to take into account various trade-offs in the type of models we chose to fit. For example, some methods take more information about the user and the song into account (e.g. metadata) while others use a collaborative filtering approach without taking into account user or song characteristic such as age, genre, location, etc.
There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.

For our project, we explored various song recommendation systems algorithms such as popularity recommendations, artist based recommensations, matrix factorization, and user and item based similarity measures.  
% memory based collaborative filtering

@@

\subsection{User Based and Item Based Collaborative Filtering}
%\subsubsection{Intro to Memory-Based Collaborative Filtering}
A simpler set of methods for determining which songs to recommend to individual users is by using user- or item/song- based similarity (both also considered collaborative filtering). These only consider whether or not a user listened to a song and does not take into account the play count. Similarity scores between users (i.e. user-based) and songs (i.e. item-based) were calculated using cosine similarity. Intuitively, they work by recommending users songs that other similar users like (user-based) or by recommending songs which are similar to the songs the user has already listened to (item-based).

\subsubsection{User-Based Collaborative Filtering}
For user-based recommendation, we first calculated the similarity score between every pair of users, $u$ and $v$, using the equation below:

@@

\subsubsection{Item-Based Collaborative Filtering}
For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, $i$ and $j$, using the equation below and saved the most similar songs to each song in the process:

@@

\section{Conclusion}
After comparing the various algorithms, the simpler models gave us the best results. Our most complicated model, matrix factorization, gave us results even below the popularity baseline. On the other hand simply recommending users songs by artists they have already listened to (the artist based baseline) resulted in a significantly better recommendation system. Lastly, our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization. Our final results for each technique employed are summarized in the table below.

\begin{table}[h]
\begin{center}
\begin{tabular}{lc}

\hline
\bf{Algorithm} &  \bf{MAP score}\\ \hline
Artist Based  Baseline  & 0.0448    \\ 
%\hline
Top 500 Songs by Count  Baseline &  0.0138  \\ 
%\hline
Top 500 Songs by Plays  Baseline &  0.0126  \\ 
%\hline
User Based  CF  &  0.0377 \\ 
%\hline
Item Based  CF  &  0.0479 \\ 
%\hline
Matrix Factorization  &   0.0143  \\ 
%\hline
\end{tabular}
\end{center}
\end{table}
